---
title: 'A Beam-Beam Collision'
teaching: 10
exercises: 20
---

:::::::::::::::::::::::::::::::::::::: questions 

How to use WarpX for beam-beam simulations of colliders 💥?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

Learn how to setup, run, and visualize your own beam-beam ☄️☄️ simulation

::::::::::::::::::::::::::::::::::::::::::::::::

![A snapshot of a beam-beam simulation.](https://gist.github.com/user-attachments/assets/217dfcde-8a77-4f73-b4bb-26c83e921255){alt="snapshot of a beam-beam simulation"}


## Setup

In this example we will simulate a bunch of electrons colliding against a bunch of positrons.
We have selected the parameters of the C$^3$ linear collider, arbitrarily. 

Some notable details:

*  Poisson solver 

*  Resolution

*  Timestep 

*  QED lookup tables

*  Diagnostics


Whenever you need to prepare an input file, [this is where you want to go](https://warpx.readthedocs.io/en/latest/usage/parameters.html).

:::::::::::::::::::::::::::::::::::::::::: spoiler

### Let's take a look at the input file

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/inputs_3d_beambeam_C3.txt
```
::::::::::::::::::::::::::::::::::::::::::::::::::

## Run 

::: caution

Coming soon. For now, [refer to the main documentation][warpx-readthedocs].
:::


## Visualize

### With Python 🐍

Now that we have the results, we can analyze them using Python.  
We will use the [openPMD-viewer][openpmd-viewer] library to grab the data that the simulation produced in `openPMD` format. 
Here you can find [a few great tutorials on how to use the viewer](https://openpmd-viewer.readthedocs.io/en/latest/tutorials/tutorials.html).
If you feel nerdy and/or you need to deal with the data in parallel workflows, you can use the [openPMD-api][openpmd-api].

As an example for the magnetic bottle simulation, we have developed simple Jupyter notebook where we retrieve the magnetic field 
and the particle attributes at the end of the simulation.
With a little bit more work, we also plot the trajectories of the particles.

:::::::::::::::::::::::::::::::::::::::::: spoiler

## Let's take a look at the Jupyter notebook

```{r, results='asis', echo=FALSE}
cat('<iframe src="https://nbviewer.org/github/aeriforme/warpx-tutorials/blob/main/episodes/files/analysis_3d_magnetic_mirror.ipynb" 
           width="100%" height="800" style="border:none;">
     </iframe>')
```
::::::::::::::::::::::::::::::::::::::::::::::::::


### With Paraview

::: caution

Coming soon!
:::


::::::::::::::::::::::::::::::::::::: keypoints 

 💅 There are several details one needs to take care when setting up a beam-beam simulation

 🔍 The [**documentation**][warpx-readthedocs] is the first place to look for answers, 
 otherwise check out our [**issues**][warpx-issues] and [**discussions**][warpx-discussions] and ask there.  

 📷 To analyze and visualize the simulation results in [**openPMD**][openpmd] format, 
 you can use the [**openPMD-viewer**][openpmd-viewer] library for Python.

::::::::::::::::::::::::::::::::::::::::::::::::
