---
title: 'A Beam-Beam Collision'
teaching: 10
exercises: 20
---

:::::::::::::::::::::::::::::::::::::: questions 

How to use WarpX for beam-beam simulations of colliders 💥?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

Learn how to setup, run, and visualize your own beam-beam ☄️☄️ simulation

::::::::::::::::::::::::::::::::::::::::::::::::

![A snapshot of a beam-beam simulation.](https://gist.github.com/user-attachments/assets/217dfcde-8a77-4f73-b4bb-26c83e921255){alt="snapshot of a beam-beam simulation"}


## Setup

In this example we will simulate a bunch of electrons colliding against a bunch of positrons.
We have selected the parameters of the C$^3$ linear collider, arbitrarily. 

Whenever you need to prepare an input file, [this is where you want to go](https://warpx.readthedocs.io/en/latest/usage/parameters.html).

:::::::::::::::::::::::::::::::::::::::::: spoiler

### Let's take a look at the input file

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/inputs_3d_beambeam_C3.txt
```
::::::::::::::::::::::::::::::::::::::::::::::::::


Some notable details:

*  Poisson solver: because the beams are ultra-relativistic (125 GeV electrons and positrons) and flat, 
we use an FFT-based electrostatic solver. 
Specifically, it solves many 2D Poisson equations in the $(x,y)$ plane for each $z$.
The full 3D version of this solver is also available with `warpx.use_2d_slices_fft_solver = 0`. 

*  Resolution: the number of grid cells is reduced to fit in a laptop. For production simulations, make sure you increase the resolution.

*  Timestep: since the beams travel at the speed of light along $z$, it makes sense to choose $dt = 
However this is not strictly necessary. Sometimes it can be useful to save resources by choosing a larger timestep.
Just make sure you resolve ''well-enough'' the shortest timescale that you're interested in. 

*  QED lookup tables: here we use the built-in ones for convenience. For production runs, make sure to use tables with enough points and set the 
ranges of the $\chi$ parameter to what you need. 

*  Diagnostics:
    * the trajectories of the beam particles. This diagnostic can easily take up too much memory. 
    For simulations with many macroparticles, consider using a field diagnostic. 
    * all the particles that exit the domain (`BoundaryScraping`)
    * the differential luminosity of every pair of left-ward and right-ward moving species 
    

## Run 

First things first. 
Create a new folder where you copy the [input file](./files/inputs_3d_beambeam_C3.txt). 
The simulation in small enough that you should be able to run it in serial with the Conda installation of WarpX.

::: spoiler

## Let's run in serial

```bash
warpx.3d inputs_3d_beambeam_C3.txt
```

Just like that! 💃
Note that with Conda's WarpX you can run this anywhere in your filesystem (provided that you copied there the input of course)
because WarpX's executable are in your `$PATH`.
:::



If you want to make the simulation faster and/or bigger, then you should run with the parallel version of WarpX.
The optimal setup to run the simulation depends on your hardware. 
This is an example that should work on many common laptops, even though it might not be ideal.


::: spoiler

## Let's run in parallel

This is just one way of doing it!

```bash
export OMP_NUM_THREADS=2
mpirun -np 4 <path/to/your/build/bin/warpx.3d> inputs_3d_beambeam_C3.txt 
```
:::

::: testimonial
On my laptop's CPU (12th Gen Intel® Core™ i9-12900H × 20) the serial simulation took ~195 seconds, while the parallel one ~44 seconds on 4 cores!

:::


## Visualize


### With Python 🐍
12th Gen Intel® Core™ i9-12900H × 20
Now that we have the results, we can analyze them using Python.  
We will use the [openPMD-viewer][openpmd-viewer] library to grab the data that the simulation produced in `openPMD` format. 
Here you can find [a few great tutorials on how to use the viewer](https://openpmd-viewer.readthedocs.io/en/latest/tutorials/tutorials.html).
If you feel nerdy and/or you need to deal with the data in parallel workflows, you can use the [openPMD-api][openpmd-api].

As an example for the beam-beam simulation, we have developed simple Jupyter notebook where we retrieve 
the beams' particles positions and project them on the $(z,x)$ and $(z,y)$ planes.

:::::::::::::::::::::::::::::::::::::::::: spoiler

## Let's take a look at the Jupyter notebook

```{r, results='asis', echo=FALSE}
cat('<iframe src="https://nbviewer.org/github/BLAST-WarpX/warpx-tutorials/blob/main/episodes/files/analysis_3d_beambeam.ipynb" 
           width="100%" height="800" style="border:none;">
     </iframe>')
```
::::::::::::::::::::::::::::::::::::::::::::::::::

You can [download the notebook](./files/analysis_3d_beambeam.ipynb) and try it yourself.
Remember to either run the notebook from the simulation directory or change the corresponding path in the notebook.


### With Paraview

::: caution

Coming soon!
:::


::::::::::::::::::::::::::::::::::::: keypoints 

 💅 There are several details one needs to take care when setting up a beam-beam simulation

 🔍 The [**documentation**][warpx-readthedocs] is the first place to look for answers, 
 otherwise check out our [**issues**][warpx-issues] and [**discussions**][warpx-discussions] and ask there.  

 📷 To analyze and visualize the simulation results in [**openPMD**][openpmd] format, 
 you can use the [**openPMD-viewer**][openpmd-viewer] library for Python.

::::::::::::::::::::::::::::::::::::::::::::::::
