---
title: 'LLNL HPC Innovation Center 2026: LWFA to Beamline Tutorial'
author: 'Arianna Formenti (LBNL)'
teaching: 0
exercises: 0
---

:::::::::::::::::::::::::::::::::::::: questions

- How does a simple 1D kinetic instability let you check that a
  simulate-then-analyze pipeline works, before committing to a heavier 3D
  run?
- How can a laser-wakefield accelerator stage be simulated with WarpX using a
  raw input file instead of PICMI?
- How is a self-injected electron bunch handed off from a particle-in-cell
  code to a beam-dynamics code?
- How can that bunch be transported through a real accelerator beamline with
  ImpactX?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Set up the software environment shared by all three exercises.
- Run and visualize a cheap 1D two-stream instability warm-up with WarpX.
- Simulate a self-injected, laser-wakefield accelerated electron bunch with
  WarpX, loading a raw input file from Python.
- Convert that bunch from WarpX's lab-frame output into the coordinate
  convention ImpactX expects.
- Track the bunch through the real HTU beamline lattice with ImpactX.

::::::::::::::::::::::::::::::::::::::::::::::::

## Overview

*Lesson author: Arianna Formenti (LBNL).*

This tutorial starts with a cheap warm-up, then follows one electron bunch
across two different simulation codes, mirroring how a real laser-plasma
accelerator beamline is modeled in practice.

1. **A two-stream instability warm-up** (`two_stream_instability`). A quick,
   fixed-parameter 1D WarpX run of two counter-streaming electron
   populations, used to confirm the simulate-then-analyze pipeline (WarpX,
   Python, Jupyter) works before moving on to the heavier stages below. See
   also the standalone
   [Two-Stream Instability](a-two-stream-instability.Rmd) episode, where you
   choose the physical and numerical parameters yourself instead of using
   fixed ones.
2. **A laser-wakefield acceleration (LWFA) stage** (`htu/lwfa_warpx`). WarpX
   simulates a short, intense laser pulse driving a plasma wake. A density
   downramp partway through the plasma triggers self-injection: electrons
   are trapped in the wake and accelerated to relativistic energies, forming
   a quasi-monoenergetic bunch. This stage is loosely modeled on the
   geometry of LBNL BELLA Center's Hundred Terawatt Undulator (HTU)
   beamline, at tutorial scale rather than at the scale of the real 2.5 J,
   sub-40 fs drive laser.
3. **The HTU beamline** (`htu/beamline_impactx`). ImpactX tracks that bunch
   through the real downstream beamline: a permanent-magnet quadrupole
   triplet, a bunch-compression chicane, an electromagnet quadrupole
   triplet, a spectrometer, and four undulator FODO segments, with
   `BeamMonitor` diagnostics at the same locations as HTU's real screens.

The two-stream warm-up and the LWFA stage both use raw, non-PICMI input: a
plain AMReX-style input file loaded from Python with
`pywarpx.warpx.load_inputs_file(...)`. The ImpactX stage likewise builds its
lattice directly from `impactx.elements`, the same way the
[ImpactX HTU beamline example](https://impactx.readthedocs.io/en/latest/usage/examples/htu_beamline/README.html)
does.

WarpX and ImpactX each initialize their own AMReX/MPI state, so they cannot
run in the same Python process. The handoff between them goes through a small
converter script and an intermediate file, which is also why the LWFA-to-
beamline handoff runs as three separate steps (WarpX, convert, ImpactX)
rather than one combined script.

## Installation

::::::::::::::::::::::::::::::::::::::::: instructor

If you're attending this tutorial as part of the LLNL HPC Innovation Center
session, you were likely handed a URL to a JupyterLab session already
running the tutorial Docker image -- open it and skip straight to
[Tutorial 1](#tutorial-1-two-stream-instability-warm-up).

:::::::::::::::::::::::::::::::::::::::::::::::::::

### Option A: the tutorial Docker image (recommended)

The image already has WarpX, ImpactX, and every Python package this lesson
needs, served as a browser-based JupyterLab session. This is how the
tutorial is presented live, typically from an AWS instance someone else has
already started for you.

To run it yourself instead:

```bash
docker pull ghcr.io/blast-warpx/warpx-tutorials/tutorial:latest
docker run --rm -p 127.0.0.1:3000:3000 ghcr.io/blast-warpx/warpx-tutorials/tutorial:latest
```

Open <http://localhost:3000/lab>, launch a **Terminal**, and go to:

```bash
cd ~/warpx-tutorials/episodes/files/llnl-hpc-2026
```

All the files linked below are already there -- no download needed. See the
[tutorial container README](https://github.com/BLAST-WarpX/warpx-tutorials/blob/main/containers/tutorial/README.md)
for GPU access and other image details.

:::::::::::::::::::::::::::::::::::::::::: spoiler

### Option B: running locally with Conda instead

Use this if you want to run the tutorial on your own machine outside of the
Docker image -- e.g. a laptop with no Docker available, or a cluster where
you'd rather use your own Conda installation.

We assume you have a Conda installation available on your machine.
Download the [Conda environment file](./files/llnl-hpc-2026/environment_setup.yml),
open a terminal in the directory containing it, and create the environment:

```bash
conda env create -f environment_setup.yml
```

Activate it with:

```bash
conda activate llnlhpc26-warpx-tutorial
```

The environment installs WarpX and ImpactX from Conda Forge, together with
the Python packages used by the simulations and the analysis notebook. This
tutorial has been tested with WarpX 26.09 and ImpactX 26.09.

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/environment_setup.yml
```

When you have finished working, deactivate the environment:

```bash
conda deactivate
```

To remove it completely at a later date, run:

```bash
conda env remove -n llnlhpc26-warpx-tutorial
```

::::::::::::::::::::::::::::::::::::::::::::::::::

## Tutorial 1: two-stream instability warm-up

### What we will simulate

Two cold-ish electron populations drift through each other at $\pm 0.1\,c$
in a 1D periodic box, on a neutralizing ion background. Small perturbations
in the charge density grow into an electrostatic wave until the beams trap
each other and the instability saturates -- the classic **two-stream
instability**, one of the most fundamental kinetic instabilities in plasma
physics.

Unlike the LWFA and beamline stages below, every physical and numerical
parameter here is already chosen for you, and the run itself takes only a
few seconds. The point of this warm-up is not the physics -- it is to
confirm, cheaply, that the full simulate-then-analyze pipeline (WarpX from
Python, then a Jupyter notebook reading its output) works on your machine
before you commit to the much heavier 3D LWFA run. If you want to explore
the physics of this instability in more depth, including choosing these
parameters yourself, see the standalone
[Two-Stream Instability](a-two-stream-instability.Rmd) episode.

### Download the files

If you are running from the tutorial Docker image, this is already present
at `~/warpx-tutorials/episodes/files/llnl-hpc-2026/two_stream_instability/`
-- `cd` there and skip to [Read the input file](#read-the-input-file).

Otherwise, create a `two_stream_instability` directory and place the
following files in it:

- [WarpX input file](./files/llnl-hpc-2026/two_stream_instability/two_stream_instability_input.txt)
- [Python driver script](./files/llnl-hpc-2026/two_stream_instability/run_two_stream_instability.py)

### Read the input file

Just like the LWFA stage below, this input file is loaded from Python
rather than run directly from the command line:

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/two_stream_instability/run_two_stream_instability.py
```

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the WarpX input file

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/two_stream_instability/two_stream_instability_input.txt
```

::::::::::::::::::::::::::::::::::::::::::::::::::

Two electron species (`ele1`, `ele2`) are initialized with equal density and
temperature, but opposite drift velocities $\pm 0.1\,c$ along $z$. There is
no ion background in the input file: since WarpX is not asked to solve
Poisson's equation at initialization, the fields start at zero and the
uniform electron charge density never sources a DC field, which has the
same effect as a neutralizing background. Reduced diagnostics
(`FieldEnergy`, `FieldMaximum`) record the field energy at every timestep,
so the notebook below can plot the instability's growth without touching
the full particle data.

### Run WarpX

```bash
python run_two_stream_instability.py
```

(If you are running locally with Conda, activate the environment first:
`conda activate llnlhpc26-warpx-tutorial`.)

This is a 1D simulation, so it finishes in seconds even in serial. When it
is done, your directory should contain a `diags/` subfolder with the
particle and reduced diagnostics.

### Analyze the results

If you are running from the tutorial Docker image, the notebook is already
present at
`~/warpx-tutorials/episodes/files/llnl-hpc-2026/two_stream_instability/two_stream_instability_plots.ipynb`.
Otherwise, download the
[analysis notebook](./files/llnl-hpc-2026/two_stream_instability/two_stream_instability_plots.ipynb)
into the `two_stream_instability` directory, alongside the input file and
driver script.

Open it with Jupyter:

```bash
jupyter lab two_stream_instability_plots.ipynb
```

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the analysis notebook

<iframe
  src="https://nbviewer.org/github/BLAST-WarpX/warpx-tutorials/blob/main/episodes/files/llnl-hpc-2026/two_stream_instability/two_stream_instability_plots.ipynb"
  width="100%"
  height="800"
  style="border: none"
  loading="lazy">
</iframe>

::::::::::::::::::::::::::::::::::::::::::::::::::

The notebook plots the particle phase space $(z, u_z)$ at the last
diagnostic dump, where the two beams should show up as the characteristic
"cat-eye" vortices of a saturated two-stream instability, and the electric
and magnetic field energy on a semi-log scale, where the electric field
energy should show clear exponential growth before saturating.

If both plots render without errors, your environment is ready for the
LWFA stage next.

## Tutorial 2: LWFA stage with WarpX

### What we will simulate

A short, intense laser pulse propagates into a plasma and drives a nonlinear
wake behind it. The plasma density starts low, ramps up to an injection
density, and then drops sharply partway through the target. That downramp
lowers the wake's phase velocity locally, which is enough to trap a bunch of
background plasma electrons that would otherwise just oscillate in the wake.
The trapped bunch is then accelerated over the remaining, lower-density
plateau, emerging as a quasi-monoenergetic beam of several tens of MeV.

This is the same physical mechanism used at the front end of the real HTU
beamline, but at tutorial scale: a coarser mesh and a higher plasma density
(and correspondingly shorter dephasing length) than the real experiment, so
the whole run finishes in a few minutes on a laptop.

### Download the files

If you are running from the tutorial Docker image, this is already present
at `~/warpx-tutorials/episodes/files/llnl-hpc-2026/htu/lwfa_warpx/` -- `cd`
there and skip to [Read the input file](#read-the-input-file-1).

Otherwise, create an `lwfa_warpx` directory and place the following files in
it:

- [WarpX input file](./files/llnl-hpc-2026/htu/lwfa_warpx/lwfa_warpx_input.txt)
- [Python driver script](./files/llnl-hpc-2026/htu/lwfa_warpx/run_lwfa_warpx.py)
- [Analysis utility module](./files/llnl-hpc-2026/htu/lwfa_warpx/lwfa_warpx_utils.py)

### Read the input file

Unlike the FCC-ee tutorial's `warpx.3d` command-line usage, this input file is
loaded from Python:

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/htu/lwfa_warpx/run_lwfa_warpx.py
```

`pywarpx.warpx.load_inputs_file(...)` reads a plain AMReX-style input file --
exactly the same format used on the command line -- so the physics setup
below is not PICMI. It is the same declarative input file style used
throughout the rest of this lesson, just initialized from a Python script
instead of directly from the `warpx.3d` executable. This is useful whenever
you want a Python driver around an otherwise unchanged input file, for
example to loop over parameter scans without duplicating the input file
itself.

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the WarpX input file

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/htu/lwfa_warpx/lwfa_warpx_input.txt
```

::::::::::::::::::::::::::::::::::::::::::::::::::

A few details worth calling out:

- All physical parameters are defined under `my_constants`, and the laser and
  plasma parameters are grouped and commented separately, so you can retune
  the laser intensity, spot size, or density profile without hunting through
  the rest of the file.
- The plasma density profile, `electrons.density_function(x,y,z)`, combines
  three sigmoid transitions in one expression: an entrance ramp from vacuum,
  the downramp injection transition, and an exit ramp back to vacuum. AMReX's
  inline parser does not support line continuation inside an expression
  string, so this has to stay on a single line.
- `warpx.do_moving_window = 1` keeps the simulation box following the laser
  pulse, so the mesh only needs to be as long as the pulse and the
  accelerating bucket, not the full plasma target.
- The diagnostic is restricted to the `electrons` species with
  `diag1.species = electrons`; there is only one species in this simulation,
  but this is the same pattern you would use to select specific species out
  of a simulation with several.

### Run WarpX

```bash
python run_lwfa_warpx.py
```

(If you are running locally with Conda, activate the environment first:
`conda activate llnlhpc26-warpx-tutorial`.)

When the run finishes, your directory should contain:

```text
lwfa_warpx_input.txt
run_lwfa_warpx.py
lwfa_warpx_utils.py
diags/
warpx_used_inputs.txt
```

## Tutorial 3: HTU beamline with ImpactX

### Download the files

If you are running from the tutorial Docker image, this is already present
at `~/warpx-tutorials/episodes/files/llnl-hpc-2026/htu/beamline_impactx/` --
`cd` there and skip to
[From WarpX's lab frame to ImpactX's beamline frame](#from-warpxs-lab-frame-to-impactxs-beamline-frame).

Otherwise, create a `beamline_impactx` directory (as a sibling of the
`lwfa_warpx` directory, so the relative path below finds the WarpX output)
and place the following files in it:

- [Converter script](./files/llnl-hpc-2026/htu/beamline_impactx/warpx_to_beamline_impactx.py)
- [Coordinate transform utilities](./files/llnl-hpc-2026/htu/beamline_impactx/transformation_utilities.py)
- [HTU lattice definition](./files/llnl-hpc-2026/htu/beamline_impactx/htu_lattice.py)
- [Python driver script](./files/llnl-hpc-2026/htu/beamline_impactx/run_beamline_impactx.py)

### From WarpX's lab frame to ImpactX's beamline frame

WarpX advances every particle at a fixed time, in lab-frame coordinates.
ImpactX instead advances a bunch through a lattice at a fixed longitudinal
position *s*, relative to a reference particle. Handing a WarpX bunch to
ImpactX therefore needs two things: a way to pick the reference particle, and
a coordinate transform between the two conventions.

[`warpx_to_beamline_impactx.py`](./files/llnl-hpc-2026/htu/beamline_impactx/warpx_to_beamline_impactx.py)
does both:

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/htu/beamline_impactx/warpx_to_beamline_impactx.py
```

1. It reads the `electrons` species from the last WarpX diagnostic dump.
2. It applies a kinetic-energy cut (`--energy-cut-MeV`, 20 MeV by default) to
   separate the accelerated bunch from the cold background plasma electrons,
   which never leave the eV range.
3. It defines the reference particle as the mean position and momentum of the
   selected bunch.
4. It transforms every selected particle's coordinates from WarpX's lab-frame,
   fixed-time convention to ImpactX's reference-relative, fixed-*s*
   convention, using
   [`transformation_utilities.py`](./files/llnl-hpc-2026/htu/beamline_impactx/transformation_utilities.py)
   -- the same coordinate-transform utility ImpactX ships with its
   [`initialize_from_array`](https://impactx.readthedocs.io/en/latest/usage/examples/initialize_from_array/README.html)
   example, for exactly this kind of external-code handoff.

Run it against the `lwfa_warpx` output:

```bash
python warpx_to_beamline_impactx.py ../lwfa_warpx/diags/diag1 warpx_bunch.npz
```

This prints the number of macroparticles selected, the reference kinetic
energy, and the bunch charge, and writes `warpx_bunch.npz` for ImpactX to
read.

::: caution

If you see `RuntimeError: No electrons found above the ... MeV energy cut`,
either the WarpX run did not self-inject (check the energy spectrum in the
analysis notebook first) or the energy cut is set above the bunch's actual
energy. Lower `--energy-cut-MeV` and check the spectrum again before assuming
the WarpX run itself needs retuning.

:::

### The HTU beamline lattice

[`htu_lattice.py`](./files/llnl-hpc-2026/htu/beamline_impactx/htu_lattice.py)
defines the full real HTU beamline as a sequence of `impactx.elements`: the
PMQ triplet immediately after the plasma target, drifts to a chicane used
for bunch compression, an EMQ triplet, a spectrometer dipole, and four
undulator FODO segments built from `VQ1`-`VQ8`. `BeamMonitor` elements are
placed at the same locations as HTU's real diagnostic screens (`TCPhosphor`,
`ChicaneSlit`, `Aline1`-`3`, `UC_VisaEBeam1`-`8`, and so on), so you can pull
out the beam's phase space at a specific, physically meaningful point in the
beamline.

Steering-magnet kickers (`S1`-`S4`, `VS1`-`VS8`) are present in the lattice
but held at zero current here, since there is no control-system setting to
plug in for a generic tutorial; in the real beamline these correct the orbit
between diagnostic screens.

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/htu/beamline_impactx/htu_lattice.py
```

### Run ImpactX

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/llnl-hpc-2026/htu/beamline_impactx/run_beamline_impactx.py
```

```bash
python run_beamline_impactx.py warpx_bunch.npz
```

When the run finishes, your directory should contain:

```text
warpx_bunch.npz
diags/
    reduced_beam_characteristics.0
    openPMD/
        monitor.h5
        TCPhosphor.h5
        ...
        UC_VisaEBeam8.h5
```

`diags/reduced_beam_characteristics.*` records beam moments (size, emittance)
at every lattice element. Each `BeamMonitor` additionally writes a full
phase-space dump named after it under `diags/openPMD/`.

## Analyze the results

If you are running from the tutorial Docker image, the notebook is already
present at `~/warpx-tutorials/episodes/files/llnl-hpc-2026/htu/htu_plots.ipynb`.
Otherwise, download the
[analysis notebook](./files/llnl-hpc-2026/htu/htu_plots.ipynb) into the `htu`
directory, alongside (not inside) the `lwfa_warpx` and `beamline_impactx`
directories you created above.

Open it with Jupyter:

```bash
jupyter lab htu_plots.ipynb
```

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the analysis notebook

<iframe
  src="https://nbviewer.org/github/BLAST-WarpX/warpx-tutorials/blob/main/episodes/files/llnl-hpc-2026/htu/htu_plots.ipynb"
  width="100%"
  height="800"
  style="border: none"
  loading="lazy">
</iframe>

::::::::::::::::::::::::::::::::::::::::::::::::::

The notebook first plots the WarpX electron energy spectrum and isolates the
self-injected bunch with the same energy cut used by
`warpx_to_beamline_impactx.py`, reporting its charge, mean energy, and energy
spread. It then reads the ImpactX reduced beam characteristics to show how
the beam size and emittance evolve along the HTU lattice, and compares the
transverse phase space at the beamline's entrance monitor against the last
undulator screen, `UC_VisaEBeam8`.

::::::::::::::::::::::::::::::::::::: keypoints

- A cheap, fixed-parameter 1D simulation is a good way to verify a
  simulate-then-analyze pipeline before committing to a heavier 3D run.
- A raw AMReX-style input file can be loaded from Python with
  `pywarpx.warpx.load_inputs_file(...)`, without going through PICMI.
- A density downramp is a simple, deterministic way to trigger self-injection
  in an LWFA simulation, without needing a separate artificial beam species.
- WarpX and ImpactX cannot share a Python process, since each owns its own
  AMReX/MPI state; coupling them means writing an intermediate file and
  running each as its own step.
- Handing a particle bunch from a time-based PIC code to a position-based
  beam-dynamics code requires both a reference-particle choice and an
  explicit fixed-time-to-fixed-*s* coordinate transform.
- A lattice built from real accelerator hardware parameters (magnet
  strengths, drift lengths, screen locations) can be tracked in ImpactX the
  same way a textbook lattice can.

::::::::::::::::::::::::::::::::::::::::::::::::
