---
title: 'OSSFE 2025 - Using WarpX, a general purpose particle-in-cell code'  
teaching: 30
exercises: 30
---

:::::::::::::::::::::::::::::::::::::: questions 

- 🤌 What is WarpX? 
- 🧮 How can I install and run WarpX?
- 🕵️ How can I analyze the simulation results?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- 💻 Install WarpX on your local machine with [Conda](https://docs.conda.io/en/latest/)
- 🏃 Run a fusion-relevant example on your local machine: protons in a magnetic mirror!
- 👀 Visualize the results with `python` and `Paraview`

::::::::::::::::::::::::::::::::::::::::::::::::

## WarpX, a particle-in-cell code

Welcome to the WarpX tutorial at [OSSFE 2025](https://ossfe.github.io/)!

[WarpX](https://github.com/BLAST-WarpX/warpx) is a general purpose **open-source** **high-performance** Particle-In-Cell (PIC) code.  
If you are not familiar with the PIC method, here is a picture that condenses the core idea:  

If you want to know more, here are a few references:

*  The two bibles on PIC 📚
     *  [C. K. Birdsall and A. B. Langdon. Plasma Physics Via Computer Simulation.](https://doi.org/10.1201/9781315275048)  
     *  [R. W. Hockney and J. W. Eastwood. Computer simulation using particles.](https://doi.org/10.1201/9780367806934)  
*  An old review written by one of the pioneers: [John M. Dawson, Particle simulation of plasmas, Rev. Mod. Phys. 55, 403](https://doi.org/10.1103/RevModPhys.55.403)  
*  Browse our docs for [many more references about advanced algorithms and methods](https://warpx.readthedocs.io/en/latest/theory/pic.html).


In this tutorial we will go through the basics of WarpX: installation, running a simple example and visualizing the results. 
Along the way, we will point to some specific locations in the documentation, for your reference.


::: callout
Everything you need to know to use WarpX is in the [documentation](https://warpx.readthedocs.io/en/latest/index.html), check it out!
:::

::::::::::::::::::::::::::::::::::::::::::::::: checklist

Some cool features of WarpX:  

- [x] Open-source - we wouldn't be here otherwise  
- [x] Runs on GPUs: NVIDIA, AMD, and Intel  
- [x] Runs on multiple GPUs or CPUs, on systems ranging from laptops to supercomputers
- [x] Advanced algorithms and methods: mesh-refinement, embedded boundaries, electrostatic/electromagnetic/pseudospectral solvers, etc.
- [x] Standards: [openPMD](https://www.openpmd.org/#/start) for output data, [PICMI](https://github.com/picmi-standard/picmi) for inputs 
- [x] Active development and mainteinance: check our [GitHub repo](https://github.com/BLAST-WarpX/warpx)
- [x] International, cross-disciplinary community: plasma physics, fusion devices, laser-plasma interactions, beam physics, plasma-based acceleration, astrophysics 
 
:::::::::::::::::::::::::::::::::::::::::::::::::::::::::

## Installing WarpX using Conda-Forge

First, you need a Conda installation and we will assume that you indeed have one.  
We will also assume you have some familiarity with the terminal. 
If not, follow the [instruction at this link](https://docs.conda.io/projects/conda/en/stable/user-guide/install/index.html#regular-installation).  
You can install Conda on most operative systems: Windows, macOS, and Linux.  
Once you have Conda on your system, WarpX is available as a package via [Conda-Forge](https://conda-forge.org/download/).  
It's a one-liner, no-brainer, small-potatoes 😌!

::: callout
```bash
conda install -c conda-forge warpx 
```
:::

Ok, maybe two lines if you want to keep your system clean by creating a new environment. 

```bash
conda create -n warpx -c conda-forge warpx 
conda activate warpx 
```

Now you should have 4 different WarpX binaries in your `PATH` called `warpx.1d`, `warpx.2d`, `warpx.3d`, `warpx.rz`.  
Each binary for a different dimensionality.  

To check this, run:
```bash
which warpx.1d warpx.2d warpx.3d warpx.rz
```
If you get 3 different paths that look something like:
```bash
/home/<username>/anaconda3/envs/warpx/bin/warpx.xd
```
then you got this 🙌! You can also import `pywarpx` in `python`


## A simple example of a magnetic mirror

In this example we will simulate a bunch of protons inside a magnetic mirror machine. 
The protons are initialized with random positions and velocities. 
The magnetic field is loaded from a `.h5` file.
Make sure to download the [input file](./files/inputs_3d_magnetic_mirror.txt). 

:::::::::::::::::::::::::::::::::::::::::: spoiler

### Let's take a look at the input file

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/inputs_3d_magnetic_mirror.txt
```
::::::::::::::::::::::::::::::::::::::::::::::::::

Now that we have an idea of what the input files looks like, let's set up our environment.
Activate the `warpx` environment if you need to.
Create a new directory with your own copy of the input file. 
Also, don't forget to [download the field file](../files/example-femm-3d.h5) and place it in the directory where you will run the input. 

::::::::::::::::::::::::::::::::::::: challenge

## Let's run the code

How would you do it? 🤷

:::::::::::::::: solution

```bash
warpx.3d inputs_3d_magnetic_mirror.txt
```
:::::::::::::::::::::::::
:::::::::::::::::::::::::::::::::::::::::::::::


You should see a standard output flashing out a lot of info.  
At the end, you should find in your folder:

  * a subfolder called `diags`: here is where the code stored the diagnostics  
  * a file called `warpx_used_inputs`: this is a summary of the inputs that were used to run the simulation  

If that's the case, yey! 💯  

If the run went wrong, you may find a `Backtrace.0.0` file which can be useful for debugging purposes. 
Let me know if the code fails in any way! 



## Visualizing with python

Now that we have the results, we can analyze them using `python`.  

:::::::::::::::::::::::::::::::::::::::::: spoiler

### Let's take a look at a simple Jupyter notebook

```{r, results='asis', echo=FALSE}
cat('<iframe src="https://nbviewer.org/github/aeriforme/warpx-tutorials/blob/main/episodes/files/analysis_3d_magnetic_mirror.ipynb" 
           width="100%" height="800" style="border:none;">
     </iframe>')
```
::::::::::::::::::::::::::::::::::::::::::::::::::


## Visualizing with Paraview 

Add stuff...

You can get images like this.
![Protons trajectories in a magnetic mirror](https://gist.github.com/user-attachments/assets/24b11226-4242-4958-bc12-c09159363065){alt="simulation of proton trajectories inside a magnetic mirror"}
...and videos like this: 







::::::::::::::::::::::::::::::::::::: keypoints 

- WarpX is a open-source high-performance particle-in-cell code
- WarpX is easy to install using `Conda`: `conda -c conda-forge warpx`
- The documentation is the first place to look for answers, otherwise check out our Issues and Discussions 
- Visualizing 3D results with Paraview and `openpmd` data using `.pmd` files. 

::::::::::::::::::::::::::::::::::::::::::::::::
