---
title: 'US-FCC 2026: Beam-Beam and Tracking Tutorial'
teaching: 0
exercises: 0
---

:::::::::::::::::::::::::::::::::::::: questions

- How can WarpX, Xsuite, ImpactX, and MAD-X be used together for FCC-ee studies?
- How can a beam-beam interaction be included in a multi-turn simulation?
- How closely do Xsuite, ImpactX, and MAD-X agree for linear tracking?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Set up the software environment used by all three exercises.
- Simulate one FCC-ee bunch crossing with WarpX.
- Combine a WarpX beam-beam interaction with multi-turn Xsuite tracking.
- Compare linear tracking through an FCC-ee lattice with Xsuite, ImpactX, and MAD-X.

::::::::::::::::::::::::::::::::::::::::::::::::

## Overview

The three exercises cover different scales of an FCC-ee simulation.

1. **A single bunch crossing.** WarpX models the electromagnetic interaction
   between an electron bunch and a positron bunch, including several QED
   processes.
2. **Many turns around the ring.** Xsuite tracks the bunches around the FCC-ee
   lattice and WarpX supplies the beam-beam interaction at the collision point.
3. **A tracking comparison.** Xsuite, ImpactX, and MAD-X track the same initial
   distribution through the same linear FCC-ee lattice.

All three exercises use one Conda environment. Tutorial 1 is designed to run
on a laptop in a few minutes. Its numerical resolution is deliberately modest
and should not be used as a production setup.

## Installation

Download the [Conda environment file](./files/us-fcc-2026/environment_setup.yml),
open a terminal in the directory containing it, and create the environment:

```bash
conda env create -f environment_setup.yml
```

Activate it with:

```bash
conda activate usfcc26-warpx-tutorial
```

The environment contains WarpX, ImpactX, Xsuite, MAD-X through CPyMAD,
openPMD-viewer, Jupyter, and the Python packages used by the analysis
notebooks.

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the environment file

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/us-fcc-2026/environment_setup.yml
```

::::::::::::::::::::::::::::::::::::::::::::::::::

You can use any notebook interface. For example, start Jupyter Lab from the
exercise directory with:

```bash
jupyter lab
```

When you have finished working, deactivate the environment:

```bash
conda deactivate
```

To remove it completely at a later date, run:

```bash
conda env remove -n usfcc26-warpx-tutorial
```

## Tutorial 1: One FCC-ee bunch crossing with WarpX

### What we will simulate

At the FCC-ee Z pole, a 45.6 GeV electron bunch collides with a 45.6 GeV
positron bunch at a full crossing angle of 30 mrad. The bunches are flat, with
an rms size of about 8 micrometers horizontally and 35 nanometers vertically
at the interaction point.

WarpX computes the beam-beam fields and advances both bunches through the
collision. The same simulation also generates beamstrahlung photons,
radiative Bhabha photons, and incoherent electron-positron pairs.

### Download the files

Create a directory for Tutorial 1 and place all of the following files in it:

- [WarpX input file](./files/us-fcc-2026/tutorial_001/tutorial_001_input.txt)
- [Analysis notebook](./files/us-fcc-2026/tutorial_001/tutorial_001_plots.ipynb)
- [Analysis utility module](./files/us-fcc-2026/tutorial_001/tutorial_001_utils.py)

The notebook and utility module must remain in the same directory as the
input file.

### Read the input file

The [input file](./files/us-fcc-2026/tutorial_001/tutorial_001_input.txt) is
split into constants, numerical settings, particle species, collision
processes, and diagnostics. Values defined under `my_constants` can be reused
throughout the rest of the file.

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the WarpX input file

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/us-fcc-2026/tutorial_001/tutorial_001_input.txt
```

::::::::::::::::::::::::::::::::::::::::::::::::::

The laptop setup uses:

- `100000` macroparticles in each primary bunch
- a `64 x 64 x 128` mesh
- `128` time steps
- a two-dimensional-slices FFT Poisson solver
- the built-in quantum-synchrotron lookup table
- a fixed random seed for reproducibility

The mesh is much coarser than a production FCC-ee simulation. It is sufficient
for learning the workflow and recovering the main features of the collision.
The built-in QED table also keeps the download small but has low resolution.
A quantitative physics study would require a higher-resolution QED table,
together with mesh and macroparticle convergence scans.

The primary species and their main products are:

| Species | Role |
|:---|:---|
| `beam1`, `beam2` | Initial electron and positron bunches |
| `pho1`, `pho2` | Beamstrahlung photons emitted by the two bunches |
| `vpho1`, `vpho2` | Virtual photons used by the equivalent-photon model |
| `ele_ll`, `pos_ll` | Landau-Lifshitz pair products |
| `ele_bh`, `pos_bh` | Breit-Heitler pair products |
| `ele_bw`, `pos_bw` | Breit-Wheeler pair products |
| `pho1_bha`, `pho2_bha` | Photons from the radiative Bhabha model |
| `beam1_bha`, `beam2_bha` | Scattered primary particles from that model |

Beamstrahlung is enabled directly on each primary beam. The incoherent-pair
and radiative Bhabha channels are listed in `collisions.collision_names`.
Event multipliers increase the number of sampled rare events, while particle
weights preserve the physical yield.

Notice that the radiative Bhabha products are stored in new species. This
makes the emitted photons and scattered primary particles easy to analyze
without mixing them into the original bunch species.

### Diagnostics

The simulation writes two openPMD particle diagnostics:

- `diags/particles_in` contains snapshots of particles that are still inside
  the simulation domain
- `diags/particles_out` records particles when they cross a domain boundary

The `ColliderRelevant` and `DifferentialLuminosity` reduced diagnostics are
written to `diags/reducedfiles`.

The analysis utility module combines the particles that remain in the box
with the boundary records. For escaped particles, it reconstructs their
positions at the final simulation time from their scrape time and momentum.

### Run WarpX

Activate the tutorial environment, change to the directory containing the
three downloaded files, and run:

```bash
conda activate usfcc26-warpx-tutorial
warpx.3d tutorial_001_input.txt
```

On the laptop used to prepare this tutorial, the run takes about three to four
minutes and produces roughly 80 MB of diagnostics. Runtime will vary with the
processor and storage system.

When the run finishes, your directory should contain:

```text
tutorial_001_input.txt
tutorial_001_plots.ipynb
tutorial_001_utils.py
diags/
warpx_used_inputs.txt
```

`warpx_used_inputs.txt` is generated by WarpX. It records the fully evaluated
input and is useful when an input value is defined through an expression.

### Analyze the collision

Open [the analysis notebook](./files/us-fcc-2026/tutorial_001/tutorial_001_plots.ipynb)
with your preferred Jupyter interface and run the cells in order. The notebook
expects the `diags` directory in the current Tutorial 1 directory.

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the analysis notebook

<iframe
  src="https://nbviewer.org/github/BLAST-WarpX/warpx-tutorials/blob/main/episodes/files/us-fcc-2026/tutorial_001/tutorial_001_plots.ipynb"
  width="100%"
  height="800"
  style="border: none"
  loading="lazy">
</iframe>

::::::::::::::::::::::::::::::::::::::::::::::::::

The [utility module](./files/us-fcc-2026/tutorial_001/tutorial_001_utils.py)
contains the routines used to read particle data, reconstruct boundary
particles, integrate the hourglass luminosity, and calculate the radiative
Bhabha cross section and lifetime.

:::::::::::::::::::::::::::::::::::::::::: spoiler

### View the analysis utility module

```{r engine='bash', echo=FALSE, comment=''}
cat ./files/us-fcc-2026/tutorial_001/tutorial_001_utils.py
```

::::::::::::::::::::::::::::::::::::::::::::::::::

The notebook proceeds through the following checks.

1. It calculates the FCC-ee beam parameters and the analytical linear
   beam-beam strength.
2. It matches macroparticles between the first and last iterations and fits
   the horizontal beam-beam kick.
3. It plots five snapshots of the bunches during the collision.
4. It compares the beamstrahlung and radiative Bhabha photon spectra.
5. It extracts the simulated luminosity and compares it with an hourglass
   calculation.
6. It estimates the radiative Bhabha cross section and beam lifetime.
7. It combines the three incoherent-pair channels and plots their angular and
   spatial distributions.

With the supplied laptop input, the normalized horizontal kick slope is about
`-0.0171`. The analytical value is about `-0.0183`, giving a difference of
roughly 7 percent. This is a useful check that the field solve and particle
tracking are behaving as expected at the chosen resolution.

::::::::::::::::::::::::::::::::::::: callout

The close agreement of one observable does not establish numerical
convergence. Increase the mesh resolution and macroparticle count before
drawing quantitative conclusions from the QED yields or luminosity.

::::::::::::::::::::::::::::::::::::::::::::::::

## Tutorial 2: Multi-turn WarpX and Xsuite simulation

From the Tutorial 2 directory, run:

```bash
python ./exec_tutorial_002.py --gpu 0 --ac t --wx tutorial_002_input.txt --nm 10000 --nt 20 --jobdir .
```

The full explanation and downloadable files will be added after the coupled
simulation has been validated.

## Tutorial 3: Linear tracking comparison

This exercise will compare the same FCC-ee lattice and initial particle
coordinates in Xsuite, ImpactX, and MAD-X. Its run and analysis instructions
will be added after the three implementations have been validated.

::::::::::::::::::::::::::::::::::::: keypoints

- WarpX resolves one collision in detail and writes both particle snapshots
  and collider-specific reduced diagnostics.
- The laptop input prioritizes runtime and teaching value over numerical
  convergence.
- Keeping different QED products in separate species makes their spectra and
  yields easier to inspect.
- The analysis notebook uses openPMD-viewer to connect the simulation output
  to beam-physics observables.

::::::::::::::::::::::::::::::::::::::::::::::::
